---
title: "PHP 2550: Assignment 1"
author: "Blain Morin"
date: "September 13, 2018"
output: 
  html_document:
    theme: "flatly"
---

```{r, include=FALSE}

### Load packages
library(knitr)
library(dplyr)
library(readr)
library(sjPlot)
library(kableExtra)
library(tidyr)
library(ggplot2)
library(gmodels)
library(grid)
library(gridExtra)

```


# Problem #1

**First we construct a data set with just the first observation for each individual:**

```{r, message=FALSE, warning=FALSE}

### Read in data
mcalindon = read_csv("McAlindon_Big.csv")

```

```{r}

### Get only the first observation for each id
first.ob = mcalindon %>%
  group_by(ID) %>%
  slice(1) %>%
  ungroup()

```

### a.) Summarize the average pain score for each of the 7 days of the study.

```{r}

### Take the mean of each pain column and put it in the table "ave.pains"
ave.pains = first.ob %>%
  summarise(mean.p1 = mean(pain.1, na.rm = TRUE),
            mean.p2 = mean(pain.2, na.rm = TRUE),
            mean.p3 = mean(pain.3, na.rm = TRUE),
            mean.p4 = mean(pain.4, na.rm = TRUE),
            mean.p5 = mean(pain.5, na.rm = TRUE),
            mean.p6 = mean(pain.6, na.rm = TRUE),
            mean.p7 = mean(pain.7, na.rm = TRUE))

```

**Here are the average pain measurements for each of the seven periods:**

```{r, echo = FALSE}
kable(ave.pains) %>%
  kable_styling() %>% 
  footnote(general = "Means for each period are calculated with NA values removed.")
```

### b.) Regress each pain score on age and use the summary function to create a summary table for each regression. Then use the confint function to find the 95% confidence interval for the regression slopes and produce a table with the estimates, standard errors, p-values and confidence intervals of the 7 slopes and put these in a single table.

```{r, warning = FALSE}

### Regress each pain score on age
painreg1 = lm(pain.1 ~ age, data = first.ob)
painreg2 = lm(pain.2 ~ age, data = first.ob)
painreg3 = lm(pain.3 ~ age, data = first.ob)
painreg4 = lm(pain.4 ~ age, data = first.ob)
painreg5 = lm(pain.5 ~ age, data = first.ob)
painreg6 = lm(pain.6 ~ age, data = first.ob)
painreg7 = lm(pain.7 ~ age, data = first.ob)

```

**Here are estimates, standard errors, p-values and confidence intervals for each of the regressions:**

```{r, message=FALSE, echo=FALSE, warning=FALSE}
sjt.lm(painreg1,
       painreg2,
       painreg3,
       show.se = TRUE,
       show.r2 = FALSE)

sjt.lm(painreg4,
       painreg5,
       painreg6,
       show.se = TRUE,
       show.r2 = FALSE)

sjt.lm(painreg7,
       show.se = TRUE,
       show.r2 = FALSE)
```

### c.) For each individual fit a regression of pain on time. Summarize the slopes and intercepts produced. 

```{r, warning = FALSE, message=FALSE, fig.width=12}

get.ints = function (p1, p2, p3, p4, p5, p6, p7, d1, d2, d3, d4, d5, d6, d7) {
  
  data.reg = data.frame(pain = c(p1,p2,p3,p4,p5,p6,p7),
                        day = c(d1,d2,d3,d4,d5,d6,d7))
  
  reg = lm(pain ~ day, data = data.reg)
  return(coef(reg)[1])
  
}

get.slopes = function (p1, p2, p3, p4, p5, p6, p7, d1, d2, d3, d4, d5, d6, d7) {
  
  data.reg = data.frame(pain = c(p1,p2,p3,p4,p5,p6,p7),
                        day = c(d1,d2,d3,d4,d5,d6,d7))
  
  reg = lm(pain ~ day, data = data.reg)
  return(coef(reg)[2])
  
}

indy.regs = first.ob %>% group_by(ID) %>%
  mutate(slope = get.slopes(p1 = pain.1,
                            p2 = pain.2,
                            p3 = pain.3,
                            p4 = pain.4,
                            p5 = pain.5,
                            p6 = pain.6,
                            p7 = pain.7,
                            d1 = lastdt1,
                            d2 = lastdt2,
                            d3 = lastdt3,
                            d4 = lastdt4,
                            d5 = lastdt5,
                            d6 = lastdt6,
                            d7 = lastdt7),
         ints = get.ints(p1 = pain.1,
                            p2 = pain.2,
                            p3 = pain.3,
                            p4 = pain.4,
                            p5 = pain.5,
                            p6 = pain.6,
                            p7 = pain.7,
                            d1 = lastdt1,
                            d2 = lastdt2,
                            d3 = lastdt3,
                            d4 = lastdt4,
                            d5 = lastdt5,
                            d6 = lastdt6,
                            d7 = lastdt7))

hist.slopes = indy.regs %>%
  ggplot(aes(x = slope)) + 
  geom_histogram(bins = 20) +
  geom_vline(aes(xintercept=mean(indy.regs$slope, na.rm = TRUE),color="mean"),
             linetype="dashed",
             size=1) +
  xlab("Slope Coefficient") +
  ylab("Count") +
  ggtitle("Histogram of Slope Coefficients") +
  theme_minimal() +
  scale_color_manual( labels = c(paste("Mean = ", round(mean(indy.regs$slope, na.rm = TRUE), digits = 3))), 
                      name = "Legend: ", values = c(mean = "red")) +
  theme(legend.position = "bottom")

hist.ints = indy.regs %>%
  ggplot(aes(x = ints)) + 
  geom_histogram(bins = 20) +
  geom_vline(aes(xintercept=mean(indy.regs$ints, na.rm = TRUE), color="mean"),
             linetype="dashed",
             size=1) +
  xlab("Intercept Value") +
  ylab("Count") +
  ggtitle("Histogram of Intercepts") +
  theme_minimal() +
  scale_color_manual( labels = c(paste("Mean = ", round(mean(indy.regs$ints, na.rm = TRUE), digits = 3))), 
                      name = "Legend: ", values = c(mean = "red")) +
  theme(legend.position = "bottom")

grid.arrange(hist.slopes, hist.ints, nrow = 1)

```


